---
title: "Simulating a backcross NIL population with magicsim"
output: rmarkdown::html_vignette
author: "Sarah Odell"
date: "4/11/2019"
vignette: >
  %\VignetteIndexEntry{"Simulating a backcross NIL population with magicsim"}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulating a BC6 NIL population with magic sim

The following uses magicsim to simulate 100 BC6 NIL maize lines from a cross between B73 and PT

      | B73 x PT|

      | F1  x F1 | 
      
      | F2  x B73 |
      
      | BC1  x B73 |
      
      | BC2  x B73 |
      
    |6 generations of selfing|
                        
A population of F1 lines are created, followed by 3 generations of backcrossing and multiple generactions of selfing to make Near Isogenic Lines (NILs)                 
### Installation
```{r results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
### Script to simulate 400 Double Haploid MAGIC lines

devtools::install_github('sarahodell/magicsim',force = T)

library('magicsim')
library("data.table")
library("tidyverse")
library("ggplot2")

options(scipen=999)
```

```{r}

gmap=fread('ogutmap_v4_ordered.csv',data.table=F)

# The 2 initial founder maize lines. This could be anything
founders=c("PT","B73")

### Simulate chromosome 1
c=1
### Initialize the founders
B73=indv_init(chr=1,gmap,h1_donors=c("B73"),h2_donors=c("B73"))
PT=indv_init(chr=1,gmap,h1_donors=c("PT"),h2_donors=c("PT"))

```

### Create the first round of F1s

We can create F1s from the first round of crossing.

```{r}
#Make F1s for each of the first crosses
f1=make_f1(B73,PT,gmap,chroms=1)
n=100

f2s=new("Pop",nIndv=n,indvlist=vector("list",length=n))
for(i in seq(1,n)){
  f2s@indvlist[[i]]=offspring(f1,f1,gmap,chroms=1)
}

```

### Building a Backcross 3 NIL Population.

Backcross the F1 onto B73 for 3 generations

```{r}
#Fix this so f2 is a population, not individuals
bc_pop=backcross_pop(B73,f2s,c=1,n=100,ngen=3,g_map=gmap)
```

Then selfed the backcross population for 6 generations to make NILs

```{r}
bc_nils=self_pop(bc_pop,c=1,ngen=6,g_map=gmap)
```

### Visualization

We can visualize the chromsome for 10 individuals in the population
```{r}
breaktable=make_pop_breaktable(bc_nils,n=10,c=1,het=F)
plot_Pop(breaktable,n=10,c=1,het=F)
```

Can also do for all 10 chromosomes

```{r}
B73=indv_init(chr=10,gmap,h1_donors=c("B73"),h2_donors=c("B73"))
PT=indv_init(chr=10,gmap,h1_donors=c("PT"),h2_donors=c("PT"))
f1=make_f1(B73,PT,gmap,chroms=10)
f2=offspring(f1,f1,g_map=gmap,chroms=10)

indv_breaktable=make_indv_breaktable(f2,het=F)
plot_Indv(indv_breaktable,het=F)
```
