---
title: "Simulating an F2 population with magicsim"
output: rmarkdown::html_vignette
author: "Sarah Odell"
date: "8/29/2019"
vignette: >
  %\VignetteIndexEntry{"Simulating a RIL population with magicsim"}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Simulating an F2 population with magic sim

The following uses magicsim to simulate 500 F2s from a cross between B73 and Mo17

      | B73 x Mo17|

          | F1 | 
          
          | F2 |
                        
A population of F1 lines are created, followed by 1 generation of selfing to make F2s                 
### Installation
```{r results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
### Script to simulate 400 Double Haploid MAGIC lines

#devtools::install_github('sarahodell/magicsim',force = T)

library('magicsim')
library("data.table")
library("tidyverse")
library("ggplot2")

options(scipen=999)
```

The genetic map should have the following columns:
'chr': chromosome number (integer)
'pos': physical position of marker (integer)
'scaled_cM': genetic position of marker (scaled so that the first marker is 0) (float)

```{r}
set.seed(11)
#read in your genetic map
gmap=fread('ogutmap_v4_ordered.csv',data.table=F)
gmap=gmap[,c('chr','pos','scaled_cM')]
# The 2 initial founder maize lines. This could be anything
founders=c("Mo17","B73")

### Simulate chromosomes 1 throught 10
c=10
### Initialize the founders
B73=indv_init(chr=c,gmap,h1_donors=c("B73"),h2_donors=c("B73"))
Mo17=indv_init(chr=c,gmap,h1_donors=c("Mo17"),h2_donors=c("Mo17"))

```

### Create the first round of F1s

We can create an F1 from the initial cross.

```{r}
n=500

### Create F1s from the first round of crossing
f1s=new("Pop",nIndv=n,indvlist=vector("list",length=n))

#Make F1s for each of the first crosses and add to the Pop object
for(i in seq(1,n)){
  f1s@indvlist[[i]]=make_f1(B73,Mo17,gmap,chroms=c)
}
```

### Building an F2 Population.

Self 500 F1s once

```{r}
gen=1
f2s=self_pop(f1s,c=c,ngen=gen,g_map=gmap)
```


### Visualization

We can visualize chromsome 10 for 10 individuals in the population
```{r}
pop_breaktable=make_pop_breaktable(f2s,c=10,het=T)
plot_Pop(pop_breaktable,n=10,c=10,het=T)
```

Can also do for all 10 chromosomes on one individual

```{r}
indv_breaktable=make_indv_breaktable(f2s@indvlist[[1]])

plot_Indv(indv_breaktable)
```

If you want to construct a vcf file from the breakpoints, save the table to a file
```{r}
fwrite(pop_breaktable,'F2_500_breaktable.txt',row.names=F,quote=F,sep='\t')
```